---
title: "Campaign Finance Final"
author: "Abigail Kessel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
getwd()
pacman::p_load("tidyverse","Hmisc", "skimr", "lubridate", "janitor")
```

#Import and View Data for Morgan Expenditures
```{r}
#read csv
mexp <- read_csv("Morgan_Expenditures.csv")

#make first row column names
mexp <- mexp %>%  row_to_names(row_number = 1)
#clean names
mexp <- clean_names(mexp)


#make things that need to be numeric numeric
mexp$amount1 <- as.numeric(mexp$amount1)
mexp$sum_to_date <- as.numeric(mexp$sum_to_date)

#how i originally removed nas -- solid idea good to have around
#mexp = subset(morgan_expenditures, select = -c(amount, office_sought, declaration, candidate ) )

#remove NAs
mexp <- mexp[,colSums(is.na(mexp))<nrow(mexp)]

#rename account1 to account. so annoying
mexp <- mexp %>% 
       rename("amount" = "amount1")

head(mexp)
```

#morgan donations
```{r}
mdon <- read_csv("Morgan_Donations.csv")
#make first row column names
mdon <- mdon %>%  row_to_names(row_number = 1)

#clean names
mdon <- clean_names(mdon)

#get numeric values to be numeric
mdon$sum_to_date <- as.numeric(mdon$sum_to_date)
mdon$amount <- as.numeric(mdon$amount)

#removes all columns that have all na values. from stack overflow https://stackoverflow.com/questions/2643939/remove-columns-from-dataframe-where-all-values-are-na
mdon <- mdon[,colSums(is.na(mdon))<nrow(mdon)]

mdon$full_zip <- str_sub(mdon$full_zip,1, 5)

tail(mdon)
```

#import and view Lee donation data
```{r}
ldon <- read_csv("Lee_Donations.csv")

#make first row column names
ldon <- ldon %>%  row_to_names(row_number = 1)
#clean names
ldon <- clean_names(ldon)

#remove NA
ldon <- ldon[,colSums(is.na(ldon))<nrow(ldon)]

#make numeric the money ones
ldon$sum_to_date <- as.numeric(ldon$sum_to_date)
ldon$amount <- as.numeric(ldon$amount)

#ldon$name <- replace_ame()

head(ldon)
```


#Import and View Lee Expenditures Data
```{r}
#read in csv
lexp <- read_csv("Lee_Expenditures.csv")

#make first row column names
lexp <- lexp %>%  row_to_names(row_number = 1)
#clean names
lexp <- clean_names(lexp)


#make things that need to be numeric numeric
lexp$amount1 <- as.numeric(lexp$amount1)
lexp$sum_to_date <- as.numeric(lexp$sum_to_date)

#remove NAs
lexp <- lexp[,colSums(is.na(lexp))<nrow(lexp)]

#rename account1 to account
lexp <- lexp %>%
       rename("amount" = "amount1")


head(lexp)
```


#Now that I've done 100 lines of just opening the csv -- let's get into the actual analysis!

#Donations Questions

1. How much total money for each campaign?
```{r}
ltotal <- sum(ldon$amount)
mtotal <- sum(mdon$amount)

ltotal
mtotal

mtotal - ltotal
```

2.Summary Lee Donations
```{r}
summary(ldon$amount)
#interesting how lee donations are rounded off and morgan donations include the cents 
```

2.5 Summary Morgan Donations
```{r}
summary(mdon$amount)
```

#Analysis Thus Far

Morgan raised more money than Lee by 227,275. Both raised over a million dollars -- makes me wonder if this was a particularly high. The min donation for Lee is 15, while for Morgan it's 1. Lee's median donation is $1000 - which is way higher than Morgan's at 25. Both have mean donations way higher than the mean, which signals that there are a few huge donations as outliers.

Questions Raised:
How many donations per candidate?
Is this typical?
Most common donors

3. How many donations each?
```{r}
nrow_ldon <- nrow(ldon)
nrow_mdon <- nrow(mdon)


nrow_ldon
nrow_mdon
nrow_mdon/nrow_ldon
```

7.6 times more donations to Morgan!!


4. Repeat Donors Lee
```{r}
lee_repeats <- ldon %>% group_by(name) %>%
  summarise(
    Total = n(),
    Amount = sum(amount)
  ) %>% arrange(desc(Total))
  
lee_repeats

#8 people donated twice, 2 donated 3 times 
```

4.5 Repeat Donors Morgan
```{r}
#mdon %>% count(mdon$name) %>% arrange(desc(n))

morgan_repeats <- mdon %>%
  group_by(name) %>%
  summarise(
    Total = n(),
    Amount = sum(amount)
  ) %>% arrange(desc(Total))
  
  
morgan_repeats


```

5. How much non-party money?
```{r}
l_nonparty <- ltotal - 1000405.84	
l_nonparty
# a from non party!

m_nonparty <- mtotal - 1315490.33	

m_nonparty
#250844 from non party!

l_nonparty - m_nonparty
```

Lee got 338k in donations, while Morgan got 250k

6. Who else is donating then? Lee
```{r}
lcontributions <- ldon %>% group_by(receipt_type_desc) %>%
  summarise(
    Total = n(),
    Amount = sum(amount)
  ) 
  
lcontributions
```

6.5 Morgan
```{r}
mcontributions <- mdon %>% group_by(receipt_type_desc) %>%
  summarise(
    Total = n(),
    Amount = sum(amount)
  ) 
  
mcontributions
```


#chart
```{r}
receipts<- c("Individual Contribution","Other Political Committee Contribution", "Party Contribution")
Lee <- c(225209.57, 107822.00, 1006005.84)
Morgan <- c(201347.74, 48674.13,1316240.33	)

df1 <- data.frame(Lee, Morgan, receipts)
df2 <- tidyr::pivot_longer(df1, cols=c('Lee', 'Morgan'), names_to='Candidate', 
values_to="value")
head(df2)

ggplot(df2, aes(x=receipts, y=value, fill=Candidate)) +
  expand_limits( y=c(0,1500000)) +
    geom_bar(stat='identity', position='dodge') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 10, vjust = 1, hjust = 1)) +
 labs(
    title= "Campaign Donations Categorized by Receipt Description",
    x="Receipt Type",
    y="Amount in Dollars",
    )

```



#Analysis Now:

Morgan had more party money by 300k, but 50k less PAC money. She had almost 8 times as many individual donations, yet had less individual donation money overall by almost 25k. Lee got 338k in donations, while Morgan got 250k. Morgan also has a bunch more repeat donors. Almost no one donates to Lee twice, but that's probably because they give big money upfront. 

7. What PACS are donating?
```{r}
ldon %>% filter(receipt_type_desc == "Other Political Committee Contribution")
```
This gives me strong incumbents vibes. He's got the support of a lot of different professional groups.

6.5 Morgan PACS

```{r}
mdon %>% filter(receipt_type_desc == "Other Political Committee Contribution")
```

7. What zipcode donated the most? How much money for each zipcode?
```{r}
ldon_zips <- ldon %>% group_by(full_zip) %>%
  summarise(
    Total = n(),
    Amount = sum(amount)
  ) %>% arrange(desc(Amount))
  
ldon_zips <- na.omit(ldon_zips)

ldon_zips
```


Morgan Zip Codes
```{r}
mdon_zips <- mdon %>% 
  group_by(full_zip) %>%
  summarise(
    Total = n(),
    Amount = sum(amount)
  ) %>%arrange(desc(Amount))

mdon_zips <- na.omit(mdon_zips)

mdon_zips
```

#district 7 zip codes
```{r}
d7_zipcodes<- 
c(28403,
28412,
28411,
28409,
28405,
28401,
28428,
28429,
28480,
28449,
28402,
28404,
28407,
28406,
28408,
28410)
```

How Morgan much money came from the district itself?
```{r}
mdon_district_zips <- mdon_zips %>% filter(full_zip %in% d7_zipcodes) 

mdistrictmoney<- sum(mdon_district_zips$Amount)
mdistrictmoney

mdistrictmoney/m_nonparty * 100
```

How much money isn't from the district
```{r}
m_nonparty - mdistrictmoney
(m_nonparty - mdistrictmoney)/m_nonparty * 100
```


How much Lee Money came from the district itself?
```{r}
ldon_district_zips <- ldon_zips %>% filter(full_zip %in% d7_zipcodes) 

ldistrictmoney<- sum(ldon_district_zips$Amount)
ldistrictmoney

ldistrictmoney/l_nonparty * 100

```

From outside the district?
```{r}
l_nonparty - ldistrictmoney
(l_nonparty - ldistrictmoney)/l_nonparty * 100
```


#Chart 1 -- Making a map of Donors for Morgan
```{r}
#most of the map stuff is from this website https://rpubs.com/richmajerus/508643

library(leaflet)

library(tigris)

zipcodes <- zctas(cb = TRUE, starts_with = c("27", "28"))

zipcodes

char_zips <- geo_join(zipcodes,
                      mdon_zips,
                      by_sp = "GEOID20",
                      by_df = "full_zip",
               how = "left")

pal <- colorNumeric( palette = "Greens",
  domain = char_zips$Amount)


labels <- 
  paste0(
    "Zip Code: ",
    char_zips$GEOID20, "<br/>",
    "Amount: ",
    scales::dollar(char_zips$Amount)) %>%
  lapply(htmltools::HTML)

```


```{r}

char_zips <- sf::st_transform(char_zips, crs = 4326)

mdonations_map <- leaflet(char_zips) %>% 
  # add base map
  addProviderTiles("CartoDB") %>% 
  setView(-79.055847, 35.913200, zoom = 7) %>% 
  # add zip codes
  addPolygons(fillColor = ~pal(Amount),
weight = 1,
opacity = 1,
color = "white",
fillOpacity = 0.7,
highlight = highlightOptions(weight = 2,
                             color = "#666",
                             fillOpacity = 0.7,
                             bringToFront = TRUE), 
label = labels) %>%
addLegend(pal = pal,
          values = ~Amount,
          opacity = 0.7,
          title = htmltools::HTML("Amount Donated per Zip Code"),
          position = "bottomright")


mdonations_map
```




#Making a map for lee donations
```{r}
lee_char_zips <- geo_join(zipcodes,
                      ldon_zips,
                      by_sp = "GEOID20",
                      by_df = "full_zip",
               how = "left")

pal2 <- colorNumeric( palette = "Greens",
  domain = lee_char_zips$Amount)


labels2 <- 
  paste0(
    "Zip Code: ",
    char_zips$GEOID20, "<br/>",
    "Amount: ",
    scales::dollar(lee_char_zips$Amount)) %>%
  lapply(htmltools::HTML)
```

```{r}
lee_char_zips <- sf::st_transform(lee_char_zips, crs = 4326)

ldonations_map <- leaflet(lee_char_zips) %>% 
  # add base map
  addProviderTiles("CartoDB") %>% 
  setView(-79.055847, 35.913200, zoom = 7) %>% 
  # add zip codes
  addPolygons(fillColor = ~pal2(Amount),
weight = 1,
opacity = 1,
color = "white",
fillOpacity = 0.7,
highlight = highlightOptions(weight = 2,
                             color = "#666",
                             fillOpacity = 0.7,
                             bringToFront = TRUE), 
label = labels2) %>%
addLegend(pal = pal2,
          values = ~Amount,
          opacity = 0.7,
          title = htmltools::HTML("Amount Donated per Zip Code"),
          position = "bottomright")


ldonations_map
```




#Pivoting to Expenditures 

#i didn't end up using any of this but i thought i could catch them doing some sneaky stuff. they didnt. just spent a shit ton on marketing

1. How much money did each campaign spend?
```{r}
sum(lexp$amount)
sum(mexp$amount)
```

Morgan spent about 13k more than Lee. She had more money after all

2. What was that money spent on?
```{r}
lpurposes <- lexp %>% group_by(purpose) %>%
  summarise(
    #Total = n(),
    Amount = sum(amount)
  ) %>% arrange(desc(Amount))

lpurposes 
```
You can just NA your operating expenses? what the heck!!! -- after further research he gave everything a purpose code so I guess that's ok??


2.5 Who exactly is Lee giving money to?
```{r}
lexp %>% group_by(name) %>%
  summarise(
    Total = n(),
    Amount = sum(amount)
  ) %>% arrange(desc(Amount))
```

Checking to see what NA purpose purchases Lee is making
```{r}
lexp %>% filter(is.na(purpose))
```
"E	Electioneering communication	Groups (other than PACs) making electioneering communications"


Every purchase has a Purpose Type Code, so I guess that balances out 
```{r}
lexp %>% group_by(purpose_type_code) %>%
  summarise(
    Total = n(),
    Amount = sum(amount)
  ) %>% arrange(desc(Amount))
```
3. Morgan Expenditures Purpose
```{r}
mexp %>% group_by(purpose) %>%
  summarise(
    Total = n(),
    Amount = sum(amount)
  ) %>% arrange(desc(Amount))
```

```{r}
mexp %>% group_by(name) %>%
  summarise(
    Total = n(),
    Amount = sum(amount)
  ) %>% arrange(desc(Amount))
```


Morgan has no NA purposes, but has NA purpose codes. I guess both parties dont want to fill out forms 
```{r}
mexp %>% group_by(purpose_type_code) %>%
  summarise(
    Total = n(),
    Amount = sum(amount)
  ) %>% arrange(desc(Amount))
```







